---
title: "Shewanellaceae in SM"
author: "André Soares"
date: "14 March 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Filtering the SM dataset for Shewanellaceae OTUs

From files in Dropbox folder "SM Paper 01 Supplements":

Table S1_79199_0_data_set_1474767_nx5v7y.xls --> renamed 'SM_LV.metadata' - **Metadata table**  
Table S2_79199_0_data_set_1474789_nx5wfh.txt --> renamed 'SM_LV.shared' - **OTU table**  
Table S3_79199_0_data_set_1474790_nx6289.txt --> renamed 'SM_LV.database' - **Taxonomical table**

# **Step 1**:
  Using 'grep' from the Linux CLI, the Taxonomical table was filtered for any **Shewanellaceae hits**.
```{r, eval=FALSE, include=TRUE}
grep "Shewanella" SM_LV.database > Shwnl_OTUs.txt
#Got 16 hits - OTUs
```
  File Shwnl_OTUs.txt now contains **.fasta sequence and RDP, SILVA and Greengenes ID's for all Shewanellaceae OTUs in the SM dataset**.
```{r, include=TRUE, eval=FALSE}
head -1 Shwnl_OTUs.txt

Otu002008	SD11SED.1020180_8075	T--AC--GA-GG-GGT---GCA-A-G--C-G--T--T--AA-T-CGG-AA----TT-A--C-T--GG-GC--GT-A--AA-GC-GT-AC--G-CA-G-G-C-G--G--T-TT-G-T-T-----AA--G-C-G-A---G-A-T-G--TG--A-AA-GC--C-C-CG-G-G---CT-C-AA-C-C-T-G-G-G-A--A-C-TBacteria(100);Proteobacteria(100);Gammaproteobacteria(100);Alteromonadales(99);Shewanellaceae(66);Shewanella(66);unclassified;unclassified;unclassified;	k__Bacteria(100);p__Proteobacteria(100);c__Gammaproteobacteria(100);o__Alteromonadales(79);f__Shewanellaceae(65);g__Shewanella(65);s__benthica(60);	Bacteria(100);Proteobacteria(100);Gammaproteobacteria(100);Alteromonadales(99);Shewanellaceae(68);Shewanella(68);
```

# **Step 2**: 
  Filter the SM OTU table for the OTU IDs found with **Step 1**:
  
  Extract OTU ID's from Shwnl_OTUs.txt
```{r, include=TRUE, eval=FALSE}
cut -f1 Shwnl_OTUs.txt > Shwnl_OTUs_list.txt
```

  Using Shwnl_OTUs_list.txt to filter the OTU table in R:
```{r, eval=FALSE}
setwd("~/Desktop/SM")
library(dplyr)
library(picante)
SM <- read.table("SM_LV.shared", header = TRUE, sep='\t', check.names=FALSE)
dim(SM)
#45982 OTUs in 974 sites

Shwnl_OTU_list <- read.table("Shwnl_OTUs_list.txt", sep=',')

colnames(SM)[0] <- "OTU"
colnames(Shwnl_OTU_list)[1] <- "OTU"

Shwnl_inSM <- inner_join(Shwnl_OTU_list, SM, by="OTU")

#Table still huge, eliminate samples with 0 sequences overall:

#turns 1st column into rownames
clean_Shwnl_inSM <- Shwnl_inSM[,-1]
rownames(clean_Shwnl_inSM)<-Shwnl_inSM[,1]

#any column summing 0 is eliminated
Shwnl_inSM_noSing <- clean_Shwnl_inSM[,colSums(clean_Shwnl_inSM) > 0]

#245 samples containing 1 or more sequences attributed to a Shewanellaceae OTU (n=16)
```


# **Step 3**:
  Using the Shewanellaceae-specific table to get OTU and sequence numbers per metadata columns of interest.  
  Importing SM metadata:
```{r, eval=FALSE}
Shwnl_metadata <- read.table("SM_LV.metadata", header = TRUE, sep=',', check.names=FALSE)
#937 samples described by 16 E.V.
```

  Extracting Shewanellaceae sample list from the Shewanellaceae-specific OTU table:
```{r, eval=FALSE}
Spl_list_Shwnl_inSM<-as.data.frame(colnames(Shwnl_inSM_noSing))
colnames(Spl_list_Shwnl_inSM)[1] <- "SampleID"
#245 samples
```

  Filtering metadata for samples containing Shewanellaceae OTUs:
```{r, eval=FALSE}
metadata_Shwnl_inSM <- left_join(Spl_list_Shwnl_inSM, Shwnl_metadata, by="SampleID")
#245 samples
```
  Transposing OTU table for grouped statistics:
```{r, eval=FALSE}
t_Shwnl_inSM_noSing <- t(Shwnl_inSM_noSing)
t_Shwnl_inSM_noSing <- data.frame(t_Shwnl_inSM_noSing[,-1], row.names=t_Shwnl_inSM_noSing[,1])
```

Numbers of OTUs per sponge species (column SCIE_NAME) and country:
```{r, eval=FALSE}
library(vegan)
Shwnl_OTUs_per_SpongeSpecies<-specnumber(t_Shwnl_inSM_noSing, metadata_Shwnl_inSM$SCIE_NAME)
Shwnl_OTUs_per_Country<-specnumber(t_Shwnl_inSM_noSing, metadata_Shwnl_inSM$COUNTRY)
```

Numbers of sequences per sponge species and country:
```{r, eval=FALSE}
Shwnl_Seqs_per_SpongeSpecies<-aggregate(t_Shwnl_inSM_noSing, by=list(metadata_Shwnl_inSM$SCIE_NAME), FUN=sum)
Shwnl_Seqs_per_Country<-aggregate(t_Shwnl_inSM_noSing, by=list(metadata_Shwnl_inSM$COUNTRY), FUN=sum)
```

# **Step 4**:
  Getting total OTU and sequence numbers from SM metadata table, as those are provided from the start.

```{r, eval=FALSE}
clean_metadata <- Shwnl_metadata[,-1]
rownames(clean_metadata)<-Shwnl_metadata[,1]

#Formatting table to aggregate
clean_metadata[,"SampleID"] <- NA
clean_metadata<-clean_metadata[,c(ncol(clean_metadata),1:(ncol(clean_metadata)-1))]
clean_metadata[,1] <- rownames(clean_metadata)
colnames(clean_metadata)[1]<-"SampleID"
rownames(clean_metadata) <- NULL

#Aggregating by Country and Sponge species
Country<-as.data.frame(clean_metadata[,c(8,4,5)])
SM_Country_OTUs_Seqs<-aggregate(. ~ COUNTRY, data=Country, FUN=sum)
Sponges<-as.data.frame(clean_metadata[,c(2,4,5)])
SM_Sponges_OTUs_Seqs<-aggregate(. ~ SCIE_NAME, data=Sponges, FUN=sum)

#Cutting out countries and sponge species not present in the Shwnl table
country_shwnl_list<-as.data.frame(Shwnl_Seqs_per_Country$Group.1)
colnames(country_shwnl_list)[1]<-"COUNTRY"
SM_FiltbyShwnlCountry_OTUs_Seqs <- inner_join(country_shwnl_list, SM_Country_OTUs_Seqs, by="COUNTRY")

sponge_shwnl_list<-as.data.frame(Shwnl_Seqs_per_SpongeSpecies$Group.1)
colnames(sponge_shwnl_list)[1]<-"SCIE_NAME"
SM_FiltbyShwnlSponges_OTUs_Seqs <- inner_join(sponge_shwnl_list, SM_Sponges_OTUs_Seqs, by="SCIE_NAME")

```



```{r, eval=FALSE}
#Writing to files:
write.table(Shwnl_OTUs_per_SpongeSpecies, "Shwnl_OTUs_per_SpongeSpecies.csv", sep = ",",quote=FALSE)
write.table(Shwnl_OTUs_per_Country, "Shwnl_OTUs_per_Country.csv", sep = ",",quote=FALSE)
write.table(Shwnl_Seqs_per_SpongeSpecies, "Shwnl_Seqs_per_SpongeSpecies.csv", sep = ",",quote=FALSE)
write.table(Shwnl_Seqs_per_Country, "Shwnl_Seqs_per_Country.csv", sep = ",",quote=FALSE)
write.table(SM_FiltbyShwnlSponges_OTUs_Seqs, "SM_FiltbyShwnlSponges_OTUs_Seqs.csv", sep=",", quote=FALSE)
write.table(SM_FiltbyShwnlCountry_OTUs_Seqs, "SM_FiltbyShwnlCountry_OTUs_Seqs.csv", sep=",", quote=FALSE)
```
Compiled every table in Libreoffice Calc and that was it for Shewanellaceae OTUs.

*Note: total values changed because of some previously undetected errors in sample naming. Those led R to think they were duplicates and to hide some diversity... All manually checked now :)*
